Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method

### 2023-09-01 manu-draft.qmd
finished RN edits

@fig-synapanoscmp shows the comparison of the EBASE results for the synthetic time series with noise (residual and tidal) to those from the original synthetic time series without noise. Only results for $a$ and $R$ are shown Results for both time series produced similar dissolved oxygen estimates (top-left panel, @fig-synapanoscmp), although notable differences were observed in the remaining parameters. In particular, gross primary production (P) and ecosystem respiration (R) were generally higher and more variable for results from the noisy time series, with differences in the former attributed to higher estimates in the $a$ parameter. As for the previous results, the $b$ parameter was estimated with low precision for both input time series, although results were centered around the true value of 0.251 $\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$. Gas exchange (D) was similar between the input time series, suggesting that lack of precision in the $b$ parameter had a minimal influence regardless of noise.

@@

Metabolic estimates from EBASE were further compared to those from existing methods to benchmark and evaluate potential differences from alternative model formulations used by each approach. The observed one-year continuous time series described above for water quality and weather data from Apalachicola Bay was used to compare metabolic estimates from EBASE and the Odum open-water method [@Odum56]. The Odum method used an approach described in @Murrell18 created for estuarine application that uses a gas-exchange approach from @Thebault08. For a better comparison with the EBASE methods, the gas exchange parameterization in the Odum method was modified to use the @Wanninkhof14 approach and the $b$ parameter was fixed at 0.251 $\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$ [as available in the WtRegDO R package, @Beck21]. An additional and fundamental difference between EBASE and the Odum method is that the former is likelihood based with metabolic rates estimated by fitting the model to the data, whereas the latter is strictly arithmetic based primarily on integration of the diel dissolved oxygen curve. Thus, potentially different results are expected between the approaches based on differences in formulation of the mass balance equations for metabolism, as well as alternative statistical approaches to estimating the parameters. Each comparison also evaluated differences in the metabolic estimates using the observed dissolved oxygen time series and a detided time series using weighted regression [@Beck15; @Beck21]. The latter comparison provided an assessment of metabolic rates that are expected to be minimally influenced by tidal advection, as a fundamental assumption of metabolic models using *in situ* dissolved oxygen time series from fixed monitoring stations. The prior distributions for EBASE used the default values justified in @sec-ebasemod (equations [-@eq-aprior], [-@eq-rprior], and [-@eq-bprior]). A 1 day model optimization period was used for EBASE for comparability with the Odum method that returns daily estimates.

The common set of metabolic parameters returned by the EBASE and Odum methods are net ecosystem metabolism (NEM), gross primary production (P), ecosystem respiration (R), and gas exchange (D). Pairwise comparisons of the daily estimates returned by each method were evaluated using simple summary statistics, including Pearson correlation coefficients ($\rho$) and root mean square errors (RMSE) of the linear regression fit comparing the same metabolic parameter between methods. Results were assessed separately for estimates using the observed or detided dissolved oxygen time series.

@fig-apacmpfig and `r run_reference(id = "tbl-apacmptab")` provide a comparison of the metabolic estimates from the EBASE and Odum methods. Correlations for all metabolic estimates, regardless of the input time series, were positive and significantly correlated, as expected. However, correlations were generally stronger for the metabolic results based on the detided dissolved oxygen time series. The only exception for the correlation estimates was gas exchange (D), which had a slightly weaker correlation using the detided results. @fig-apacmpfig shows that the Odum results for gas exchange were more variable than those for EBASE using the detided dissolved oxygen, where the latter often estimated gas exhange at much smaller values compared to the former. The RMSE values of each regression comparing the metabolic estimates between methods were generally lower in the detided results, excluding D which was higher with the detided dissolved oxygen relative to the observed.

The comparisons in @fig-apacmpfig and `r run_reference(id = "tbl-apacmptab")` suggest that tidal effects on continuous monitoring data can severely violate assumptions required for methods of estimating ecosystem metabolism [as described in detail in @Beck15]. These effects are easily identified by stronger correlations and lower RMSE using the detided dissolved oxygen. Further, effects of tidal advection can be seen with estimates from the Odum method as negative values for P and R, i.e., "anomalous" values. The effects of tidal advection on estimates from EBASE were more subtle as anomalous values cannot be returned based on constraints of the priors (i.e., none can be negative), although the range of values for the metabolic estimates were much larger with observed dissolved oxygen. A similar result (larger estimates for P and R) was observed using the synthetic time series with instrument and tidal noise (@fig-synapanoscmp). Overall, the results suggest that detiding the dissolved oxygen time series prior to estimating metabolic rates reduces tidal noise and produces more stable estimates with smaller ranges, regardless of the method used.

@@

The above analyses demonstrated that EBASE can successfully recover known metabolic parameters from a synthetic time series and provides robust estimates on real data covering a year of observations. The success of these results was predicated on the requirements of EBASE to address several critical needs that existing methods for estimating metabolism do not address for coastal applications. These needs were addressed by 1) using a Bayesian framework that estimates the best fit of the model to the data with posterior probabilities describing the likelihood of the fit, and 2) modifying the underlying mass-balance model for dissolved oxygen with several components more appropriate for estuarine settings. First, the Bayesian framework enables hypothesis testing of how a model can be fit to data by using prior knowledge in the model fitting process, which results in posterior probabilities describing the certainty of the parameter estimates [@Hilborn13]. The advantages of the Bayesian approach have led to widespread adoption over frequentist approaches in many scientific applications, especially in the coastal and marine community [e.g., @myers2001; @borsuk2001; @brown2017], yet these methods have not been rigorously explored for estimating metabolic rates in estuaries. Our results demonstrated that even using uninformed priors, EBASE can reproduce known metabolic parameters and rates with reasonable certainty. However, an exception was the ability of the model to estimate stable values for the $b$ parameter, although the resulting estimate of gas exchange was similar to the known values. This result could be explained by the dominance of the relative processes that manifest signals in the dissolved oxygen time series. That is, estuarine dissolved oxygen at Apalachicola Bay may be dominated by production and respiration, and less so from gas exchange, such that the available data provided insufficient information for EBASE to estimate $b$. This result aligns with others that have suggested gas exchange may be a relatively small component of the metabolic signal in shallow estuaries of the Gulf of Mexico [@Murrell18]. In such cases, more constrained prior distributions for $b$ could be appropriate or EBASE could be tested in settings with expected larger contributions of gas exchange expressed by dissolved oxygen.

The second advantage of EBASE that makes the model more appropriate for estuarine settings included several modifications to the underlying mass-balance equation for dissolved oxygen (@eq-ebase). As described in @sec-basemodif, EBASE is an estuarine adaptation of BASE [@Grace15] that was developed for stream applications and many assumptions of the underlying model of the latter were inappropriate for the former. Specifically, EBASE was modified to address three primary concerns of using BASE in an estuarine setting. First, the estimation of gross primary production as a function of PAR assumed a linear relationship such that the increase of production per unit increase of PAR was consistent for any level of the latter. The BASE model includes an exponential parameter, $p$, that defines a non-linear relationship between P and PAR (although it should be noted that BASE provides an option to set $p = 1$). @Holtgrieve10 describes a simple linear photosynthesis-irradiance curve as appropriate for systems that are never light-saturated and this assumption likely holds true for estuaries that are commonly more turbid than open-ocean or freshwater locations. Second, EBASE removed the temperature-dependency of respiration and assumed that respiration did not change with water temperature, consistent with applications of the Odum method to other shallow estuaries [@Beck15; @Murrell18]. This modification addresses recent work that suggests respiration in streams may be reduced during the night with lower temperatures as the respiration of labile autochthonous organic matter may contribute to changes in the dissolved oxygen time series [@sadro2014; @schindler2017]. There is currently no justification for this assumption in estuaries. Finally, the gas exchange in EBASE used the parameterization in @Wanninkhof14 that is primarily wind-based and developed for the open-ocean, which is clearly more appropriate than the prior formulation in BASE that includes a temperature and depth dependency for streams. Overall, these changes also reduced the required inputs and many monitoring locations nationwide include the requisite data for applying EBASE (i.e., dissolved oxygen, water temperature, salinity, PAR, and wind speed, as collected by the NERRS).

Our results also confirmed that the effects of tidal advection on metabolic estimates can lead to biases in understanding gross primary production and ecosystem respiration. This result has previously been shown using the Odum open water method [@kemp1980; @caffrey2003; @Beck15] and confirmed herein, whereby results using observed time series can produce over-estimates of metabolic rates with more variation than those without tidal influence. Our assessment of the synthetic time series with noise and use of the detided dissolved oyxgen time series from the observed data at Apalachicola Bay to estimate metabolism both demonstrated this bias (Figures [-@fig-synapanoscmp] and [-@fig-apacmpfig]). As a result, detiding dissolved oxygen data is recommended prior to estimating metabolism, regardless of whether EBASE or an alternative method is used. In addition to inaccurate and imprecise metabolic estimates, tidal influences are also most likely observed as negative (or anomalous) values in the production and respiration rates using the Odum method [@caffrey2003]. Tidal effects may be less obvious using EBASE since negative values cannot be obtained based on constraints of the prior distributions, although as noted above, larger and more variable estimates are returned. A potential future approach for developing EBASE could be an explicit formulation of tidal advection in the model using a synoptic tidal height time series in the Bayesian framework. This has the advantage of simultaneously detiding and estimating metabolic rates using the same model, whereas using weighted regression prior to EBASE, as used herein, requires additional analysis and assumptions. However, field measurements [e.g., control volume approaches, @falter2008; @nidzieko2014] may be required to provide reasonable constraints on the prior distributions for any parameters that quantify advective influences on dissolved oxygen.

EBASE provides several computational advantages over BASE implemented in the R package that was created for the purposes herein [@Beck23]. Most importantly, EBASE was developed to allow different model optimization lengths, whereas the BASE method was developed to estimate results for a single day. Allowing EBASE to incorporate more than a day of observations in the optimization can likely produce more accurate and precise estimates for the metabolic parameters. Physical and biological characteristics of an ecosystem that affect the metabolic rates are likely acting at time scales longer than a day, e.g., wind patterns that can affect gas exchange as quantified by the $b$ parameter may persist for several days. Exposing the model to additional observations may produce more stable results as the parameters for the best fit are estimated by the Bayesian routine as a function of the data that are evaluated each optimization period. However, we cannot provide a precise recommendation on the number of days to use for the optimization period as the exact length of time that ecosystem characteristics can affect metabolic rates may not be known and likely varies by location. Our example using the synthetic time series from Apalachicola Bay suggests that results were similar for the 7 and 30 day optimization period. As such, using one week of data for each optimization period is likely appropriate for this location. For other locations, using 1 day is not recommended given the issues stated above (i.e., unstable results), whereas an optimization period longer than a month may miss important events in the time series that can affect the metabolic estimates and will likely produce parameters that are overly generalized.

@@

While the EBASE theory and software provide advantages for metabolic estimates, the approach is not without limitations. Like most models, explicit choices are needed to apply EBASE to time series data. First, the use of unconstrained prior distributions was shown to reproduce known metabolic parameters with reasonable certainty, although the $b$ parameter was still relatively unstable and unconstrained priors may not provide useful estimates in estuarine settings outside of the location evaluated herein. This may require further constraining the priors, although doing so may produce biased results in the absence of additional data to justify a constraint. Field-based measurements or alternative metabolic models could be used for informing priors. As an example, initial application of BASE to estuarine data prior to developing EBASE demonstrated that unstable and unrealistic metabolic rates were returned with uninformed priors. The Odum method was used on the same data to develop an appropriate expectation of the range of values for the reaeration coefficient, $K$, which was then used to inform the prior distribution using the BASE method. More stable and realistic metabolic estimates were then returned by BASE, demonstrating the value of the Bayesian approach that can incorporate prior knowledge. A similar approach could be applied to create more accurate estimates from EBASE depending on location. An additional choice for using EBASE is the model optimization period, although some guidance is provided above.

@@

The development of EBASE represents a new approach for estimating metabolic rates in estuarine settings that leverages existing methods by adapting theoretical components of the underlying mass-balance model using a Bayesian framework. Our results have demonstrated that unknown metabolic parameters can be estimated with reasonable certainty using uninformed prior distributions and EBASE could be a valuable approach to improve the understanding of critical ecosystem processes as the method can be readily applied to other settings. A notable limitation of EBASE was an inability to precisely recover the $b$ parameter from a synthetic time series, although this result may be explained by a minimal contribution of gas exchange as expressed in the dissolved oxygen time series. Effects of tidal advection on dissolved oxygen measurements were also observed, demonstrating a violation of the assumption that the measured data represent a water sample having a continuous metabolic history. Further development of EBASE to accommodate a tidal signal, or exploring alternative methods for detiding dissolved oxygen data, are promising avenues of future research to more accurately quantify metabolic parameters in estuaries. Overall, EBASE represents a potentially powerful tool that removes many of the assumptions of existing methods by allowing unknown metabolic parameters to be estimated by the data and can also incorporate existing knowledge for further exploration of estuarine ecosystem properties.

@@

We thank Jason Garwood and Ethan Borque of the Apalachicola National Estuarine Research Reserve for assisting with the deployment and processing of data used herein. We also thank Wade McGillis for assisting in this effort. We thank Darren Giling for initial conversations on applying BASE to estuarine data. This research was funded by award 1924559 from the National Science Foundation to co-PIs Ray Najjar, Maria Herrmann, and Kathleen Hill.

>>>>

@fig-synapanoscmp shows the comparison of the results recovered from EBASE for the synthetic time series with noise (residual and tidal) and for the synthetic time series without noise. EBASE was able to recover reasonable estimates for both $a$ and $R$, although the results from the noisy time series were slightly biased towards higher values as shown by the blue regression lines on the right side of @fig-synapanoscmp. The higher values for $a$ will produce higher estimates for $P$, and combined with $R$, will produce inflated values for metabolic estimates compared to the results for the synthetic time series without noise. Although the results for the latter were not biased towards higher values, both estimates provided by EBASE had a relatively large range of variablity around the known values. As noted above, a longer optimization period could reduce this variability.

@@

Metabolic estimates from EBASE were further compared to those from existing methods to benchmark and evaluate potential differences from alternative model formulations used by each approach. The observed one-year continuous time series described above for water quality and weather data from Apalachicola Bay was used to compare metabolic estimates from EBASE and the Odum open-water method [@Odum56]. The Odum method used an approach described in @Murrell18 created for estuarine application that uses a gas-exchange approach from @Thebault08. For a better comparison with the EBASE methods, the gas exchange parameterization in the Odum method was modified to use the @Wanninkhof14 approach and the $b$ parameter was fixed at 0.251 $\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$ [as available in the WtRegDO R package, @Beck21]. An additional and fundamental difference between EBASE and the Odum method is that the former is likelihood-based with metabolic rates estimated by fitting the model to the data, whereas the latter is strictly arithmetic-based primarily on integration of the diel dissolved oxygen curve. Thus, potentially different results are expected based on the alternative statistical approaches to estimate the metabolic results. Each comparison also evaluated differences in the results using the observed dissolved oxygen time series and a detided time series using weighted regression [@Beck15; @Beck21]. The latter comparison provided an assessment of metabolic rates that are expected to be minimally influenced by tidal advection, as a fundamental assumption of metabolic models using *in situ* dissolved oxygen time series from fixed monitoring stations. The prior distributions for EBASE used the default values justified in @sec-ebasemod (equations [-@eq-aprior], [-@eq-rprior], and [-@eq-bprior]). Compared to the previous analyses, the prior distribution for the $b$ parameter was not fixed given that its estimate should vary as a function of the data using real observations. A 1 day model optimization period was used for EBASE for comparability with the Odum method that returns daily estimates.

Pairwise comparisons of the daily estimates of $NEM$, $P$, and $R$ returned by each method were evaluated using simple summary statistics, including Pearson correlation coefficients ($\rho$) and root mean square differences (RMSD) of the linear regression fit comparing the same flux type between methods. Results were assessed separately for estimates using the observed or detided dissolved oxygen time series. @fig-apacmpfig and `r run_reference(id = "tbl-apacmptab")` provide a comparison of the metabolic estimates from the EBASE and Odum methods. Correlations for all metabolic estimates, regardless of the input time series, were positive and significantly correlated, as expected. However, correlations were generally stronger (and RMSDs smaller) for the metabolic results based on the detided dissolved oxygen time series. The observed time series clearly produced a much larger range of values for results from both the EBASE and Odum methods.

The comparisons in @fig-apacmpfig and `r run_reference(id = "tbl-apacmptab")` suggest that tidal effects on continuous monitoring data can severely violate assumptions required for methods of estimating ecosystem metabolism [as described in detail in @Beck15]. These effects are easily identified by stronger correlations and lower RMSD using the detided dissolved oxygen. Further, effects of tidal advection can be seen with estimates from the Odum method as negative values for P and R, i.e., "anomalous" values. The effects of tidal advection on estimates from EBASE were more subtle as anomalous values cannot be returned based on constraints of the priors (i.e., none can be negative), although the range of values for the metabolic estimates were much larger with observed dissolved oxygen. A similar result (larger estimates for P and R) was observed using the synthetic time series with instrument and tidal noise (@fig-synapanoscmp). Overall, the results suggest that detiding the dissolved oxygen time series prior to estimating metabolic rates reduces tidal noise and produces more stable estimates with smaller ranges, regardless of the method used.

@@

The above analyses demonstrated that EBASE can successfully recover known metabolic parameters from a synthetic time series and provides robust estimates on real data covering a year of observations. The success of these results was predicated on the requirements of EBASE to address several critical needs that existing methods for estimating metabolism do not address for coastal applications. These needs were addressed orimarily using a Bayesian framework that estimates the best fit of the model to the data with posterior probabilities describing the likelihood of the fit. This framework enables hypothesis testing of how a model can be fit to data by using prior knowledge in the model fitting process, which results in posterior probabilities describing the certainty of the parameter estimates [@Hilborn13]. The advantages of the Bayesian approach have led to widespread adoption over frequentist approaches in many scientific applications, especially in the coastal and marine community [e.g., @myers2001; @borsuk2001; @brown2017], yet these methods have not been rigorously explored for estimating metabolic rates in estuaries. Our results demonstrated that even using uninformed priors, EBASE can reproduce known metabolic parameters and rates with reasonable certainty. However, initial analyses suggested an inability of the model to estimate stable values for the $b$ parameter that influence gas exchange. This parameter was fixed for the sensitivity analyses and we were unable to evaluate the ability of EBASE to recover it from the synthetic time series. Future work should focus on understanding the sensitivity of EBASE to estimating $b$ where the dominance of the relative processes that manifest signals in the dissolved oxygen time series could vary. For example, estuarine dissolved oxygen at Apalachicola Bay may be dominated by production and respiration, and less so from gas exchange, such that the available data may provide insufficient information for EBASE to estimate $b$. This aligns with others that have suggested gas exchange may be a relatively small component of the metabolic signal in shallow estuaries of the Gulf of Mexico [@Murrell18]. In such cases, more constrained prior distributions for $b$ could be appropriate or EBASE could be tested in settings with expected larger contributions of gas exchange expressed by dissolved oxygen.

Our results also confirmed that the effects of tidal advection on metabolic estimates can lead to biases in understanding gross primary production and ecosystem respiration. This result has previously been shown using the Odum open water method [@kemp1980; @caffrey2003; @Beck15] and confirmed herein, whereby results using observed time series can produce over- and under-estimates of metabolic rates with more variation than those without tidal influence (Figures [-@fig-synapanoscmp] and [-@fig-apacmpfig]). As a result, detiding dissolved oxygen data is recommended prior to estimating metabolism, regardless of whether EBASE or an alternative method is used. In addition to inaccurate and imprecise metabolic estimates, tidal influences are also most likely observed as negative (or anomalous) values in the production and respiration rates using the Odum method [@caffrey2003]. Tidal effects may be less obvious using EBASE since negative values cannot be obtained based on constraints of the prior distributions, although as noted above, more variable estimates are returned. A potential future approach for developing EBASE could be an explicit formulation of tidal advection in the model using a synoptic tidal height time series in the Bayesian framework. This has the advantage of simultaneously detiding and estimating metabolic rates using the same model, whereas using weighted regression prior to EBASE, as used herein, requires additional analysis and assumptions. However, field measurements [e.g., control volume approaches, @falter2008; @nidzieko2014] may be required to provide reasonable constraints on the prior distributions for any parameters that quantify advective influences on dissolved oxygen.

There were several considerations during the development of EBASE that makes the model appropriate for estuarine settings. As described in @sec-basemodif, EBASE is an estuarine adaptation of BASE [@Grace15] that was developed for stream applications and many assumptions of the underlying model of the latter were inappropriate for the former. The most critical change was the formulation for gas exchange in EBASE using the parameterization in @Wanninkhof14 that is primarily wind-based and developed for the open-ocean, which is clearly more appropriate than the prior formulation in BASE that includes a temperature and depth dependency for streams. EBASE also removed the temperature-dependency of respiration and assumed that respiration did not change with water temperature, consistent with applications of the Odum method to other shallow estuaries [@Beck15; @Murrell18]. Overall, these changes also reduced the required inputs and many monitoring locations nationwide include the requisite data for applying EBASE (i.e., dissolved oxygen, water temperature, salinity, PAR, and wind speed, as collected by the NERRS).

EBASE also provides several computational advantages over BASE implemented in the R package that was created for the purposes herein [@Beck23]. Most importantly, EBASE was developed to allow different model optimization lengths, whereas the BASE method was developed to estimate results for a single day. Allowing EBASE to incorporate more than a day of observations in the optimization can likely produce more accurate and precise estimates for the metabolic parameters. Physical and biological characteristics of an ecosystem that affect the metabolic rates are likely acting at time scales longer than a day; e.g., wind patterns that can affect gas exchange as quantified by the $b$ parameter (such as fetch and waves) as well as ecological and biogeochemical characteristics influencing the $a$ and $R$ parameters (such as community composition and nutrient availability) may persist for several days. Exposing the model to additional observations may produce more stable results as the parameters for the best fit are estimated by the Bayesian routine as a function of the data that are evaluated each optimization period. However, we cannot provide a precise recommendation on the number of days to use for the optimization period as the exact length of time that ecosystem characteristics can affect metabolic rates may not be known and likely varies by location. Our example using the synthetic time series from Apalachicola Bay suggested that the most precise results were obtained using the 30 day optimization period, although results are likely to vary across sytems. For most applications, very short (e.g., 1 day) or very long (e.g., \> 30 days) are not recommended, such that the former may produce unstable results and the latter may miss important events in the time series producing parameters that are overly generalized. However, an advantage of the Bayesian approach is the ability to evaluate different optimization periods and, therefore, the amount of data on which to estimate key parameters. Hypotheses can be generated on the amount of time over which critical metabolic parameters may vary across systems.

@@

While the EBASE theory and software provide advantages for metabolic estimates, the approach is not without limitations. Like most models, explicit choices are needed to apply EBASE to time series data. First, the use of unconstrained prior distributions was shown to reproduce known metabolic parameters with reasonable certainty, although initial assessments suggested the $b$ parameter was relatively unstable and unconstrained priors may not provide useful estimates. Parameters that are poorly estimated may require constrained priors, although doing so may produce biased results in the absence of additional data to justify a constraint. Field-based measurements or alternative metabolic models could be used for informing priors. As an example, initial application of BASE to estuarine data prior to developing EBASE demonstrated that unstable and unrealistic metabolic rates were returned with uninformed priors. The Odum method was used on the same data to develop an expectation of the range of values for the reaeration coefficient, $K$, which was then used to create an informed prior distribution for the same parameter using the BASE method. More stable and realistic metabolic estimates were then returned by BASE, demonstrating the value of the Bayesian approach that can incorporate prior knowledge. A similar approach using alternative metabolic models for informing priors could be applied to create more accurate estimates from EBASE. An additional choice for using EBASE is the model optimization period, although some guidance is provided above.

@@

The development of EBASE represents a new approach for estimating metabolic rates in estuarine settings that leverages existing methods in a Bayesian framework. Our results have demonstrated that unknown metabolic parameters can be estimated with reasonable certainty and EBASE could be a valuable approach to improve the understanding of critical ecosystem processes as the method can be readily applied to other settings. A notable limitation of EBASE was an inability to precisely recover the $b$ parameter from a synthetic time series, although this result may be explained by a minimal contribution of gas exchange as expressed in the dissolved oxygen time series. Effects of tidal advection on dissolved oxygen measurements were also observed, demonstrating a violation of the assumption that the measured data represent a water sample having a continuous metabolic history. Further development of EBASE to accommodate a tidal signal, or exploring alternative methods for detiding dissolved oxygen data, are promising avenues of future research to more accurately quantify metabolic parameters in estuaries. Overall, EBASE represents a potentially powerful tool that removes many of the assumptions of existing methods by allowing unknown metabolic parameters to be estimated by the data and can also incorporate existing knowledge for further exploration of estuarine ecosystem properties.

@@

We thank Jason Garwood and Ethan Borque of the Apalachicola National Estuarine Research Reserve for assisting with the deployment and processing of data used herein. We also thank Wade McGillis for assisting in this effort. We thank Darren Giling for initial conversations on applying BASE to estuarine data. This research was funded by award 1924559 from the National Science Foundation to co-PIs Raymond Najjar, Maria Herrmann, and Kathleen Hill.
